---
title: "Final Project Portfolio"
author: "Lea Frank"
date: "2/23/2019"
output:
  html_document:
    theme: spacelab
    toc: true
    toc_float: true
    code_folding: hide
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# I created a package that includes the data, in order to access you'll need to 
# install the package using the commented code below. 
# devtools::install_github("lfrank14/lurrdatviz")
# install.packages(c("ggthemes", "paletteer"))
library(lurrdatviz)
library(tidyverse)
library(ggthemes)
library(colorblindr)
library(paletteer)

```


### Data Summary

**FACAT**: An incidental category learning task. Participants studied face-name associations, with the surname indicatnig the underlying cateogry structure. Representational similarity analysis (RSA) was used to test if certain brain regions (anterior hippocampus, posterior hippocampus, and ventromedial prefrontal cortex/medial orbitofrontal cortex) were representing the unique faces (item representation) or the underlying category structure (category representations). 

**DICAT**: A feedback-based category learning task. Participants learned to categorize novel cartoon "animals" into the appropriate categories (e.g. febbles and badoons). During test, participants had to categorize old and new animals. Categorization accuracy for new animals measured the participants ability to generalize what they learned to new experiences. Behavioral models were also used to determine what strategy participants used while categorizing new animals. Under the exemplar approach, participants use previously encountered animals (exemplars) to categorize new ones. Under the prototype approach, participants use an abstract representation (prototype) to categorize new animals.

**AEPET**: An individual differences study that includes a similar task to that of DICAT, just using different cartoon animals. 

I have run some statistics on the data to communicate significant findings, but I did not include them in the html. If you are interested, you can find them in the .rmd file, directly above the associated plot. 

### Figure 1. Category and Item Representations

+ Data: FACAT 
+ Figure: Bar plot showing mean strength of item and category representations across different brain regions. 
+ Message: To show that during an incidental category learning, item representations are stronger that category representations.  
+ Audience: NIH study section
+ Notes: We only really care about the strength of each type of representation as it compares to 0. The strength of representation is measured by an effect size, Cohen's d. 

```{r facat_data}

# Labels
rois <- c("ahip","phip","mofc")
nrois <- length(rois)
measures <- c("itemrep","itemrep2","catrep")

# Load data
facat <- facat
facat_reps <- facat %>% 
  select(subnum, contains("_train_")) %>% 
  filter(subnum < 33) # missing data for most recent subjects

collabels <- colnames(facat_reps[-1])

# Tidy for plotting  
facat_tidy <- facat_reps %>% 
  gather(key = "tmp", value = "rep_strength", -1) %>% 
  separate(col = "tmp", into = c("roi", NA, "rep_type")) %>% 
  filter(rep_type != "itemrep" &
           roi != "hip")

# Summarize for plotting
facat_sum <- facat_tidy %>% 
  group_by(roi, rep_type) %>% 
  summarize(m_repstr = mean(rep_strength),
            se_repstr = se(rep_strength))
```

```{r fig1_v1_stats, inclue=FALSE}
# Compare against 0, for reference / annotations
results <- list()
for (i in 1:(ncol(facat_reps) - 1)) {
  results[[i]] <- t.test(facat_reps[i + 1])
}

pvals <- map(results, function(x)x$p.value)
sig_reps <- tibble(condition = collabels,
                   pval = pvals,
                   is_sig = pvals < .05,
                   is_marg = pvals < .1)
```


```{r fig1_v1}

ggplot(facat_sum, aes(x = factor(roi, levels = c("ahip","phip","mofc")), 
                      y = m_repstr, fill = factor(rep_type, levels = c("itemrep2","catrep")))) +
  geom_bar(stat = "identity", position = "dodge", color = "white") +
  geom_hline(yintercept = 0, color = "#343434", alpha = .5) +
  geom_errorbar(aes(ymin = m_repstr - se_repstr, ymax = m_repstr + se_repstr),
                 position = position_dodge(width = .9), width = .2, color = "#343434") +
  theme_minimal(base_size = 16) +
  #theme_light(base_size = 16) +
  #scale_fill_OkabeIto(name = NULL, labels = c("Category", "Item")) +
  scale_fill_paletteer_d(package = "rcartocolor", palette = "Vivid",
                         name = NULL, labels = c("Item", "Category")) +
  labs(y = "Cohen's d",
       x = NULL,
       title = "Strength of item and category representations during training",
       caption = "Ant.HIP = Anterior Hippocampus, Post.HIP = Posterior Hippocampus, VMPFC = Ventromedial Prefrontal Cortex") +
  scale_x_discrete(labels = c("Ant. HIP", "Post. HIP", "VMPFC")) +
  scale_y_continuous(limits = c(-.015,.035)) +
  theme(panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.major.y = element_line(size = .3),
        panel.grid.minor.y = element_line(size = .3),
        legend.position = c(.99, .99), 
        legend.justification = c("right","top"),
        axis.text.x = element_text(size = 16, color = "black"),
        axis.text.y = element_text(color = "black"),
        plot.caption = element_text(hjust = -1)) +
  annotate(geom = "text", label = "*", x = .78, y = .03, size = 10)

  # jtools::theme_apa()

# The only thing I could think of was to add color = "white" to your geom_bar line. This will add the white outline between item and category for the VMPFC. It appears that this may be something that you would use for a publication so I added the jtools package at the bottom. This pretty much creates a figure that you can include in an APA journal. I think that what you had for the caption fit better as a title, so I changed that and included a caption for the regions of the brain instead. Overall, I think this is a great simple understandable plot for a paper. The significance asterisk is a nice touch and I like the colors. The error bars are a nice touch and make for an appropriate display of uncertainty. -JP

```

```{r fig1_v2}
#facat_sum$rep_type <- factor(facat_sum$rep_type, levels = c("itemrep2","catrep"))
facat_sum$roi <- factor(facat_sum$roi, levels = c("ahip","phip","mofc"),
                        labels = c("Ant. HIP","Post. HIP","VMPFC"))

ggplot(facat_sum, aes(x = factor(rep_type, levels = c("itemrep2","catrep")), 
                      y = m_repstr, 
                      fill = factor(rep_type, levels = c("itemrep2","catrep")))) +
  geom_bar(stat = "identity") +
  geom_hline(yintercept = 0, color = "#343434", alpha = .5) +
  geom_errorbar(aes(ymin = m_repstr - se_repstr, ymax = m_repstr + se_repstr),
                width = .2, color = "#343434") +
  facet_wrap(~roi) +
  #facet_wrap(~factor(roi, levels = c("ahip","phip","mofc"),
  #                   labels = c("Ant. HIP","Post. HIP","VMPFC"))) +
  scale_fill_paletteer_d(package = "rcartocolor", palette = "Vivid",
                         name = NULL, labels = c("Item", "Category")) +
  #scale_y_continuous(limits = c(-.015, .035)) +
  #annotate(geom = "text", label = "*", x = .78, y = .03, size = 10) + # it keeps throwing me a weird error if I include this - stupid annotate doesn't work on facet_wraps very well...
  theme_minimal(base_size = 16) +
  labs(y = "Cohen's d", x = NULL,
       caption = "Figure 1. Strength of item and category representations") +
  theme(panel.grid.major.x = element_blank(),
        panel.grid.major.y = element_line(size = .3),
        panel.grid.minor.y = element_line(size = .3),
        legend.position = "bottom",
        plot.caption = element_text(hjust = -.35),
        axis.text.x = element_blank(),
        axis.text.y = element_text(color = "black")) 

# JP's recommendation
facat_annotate <- data.frame(rep_type = "itemrep2", m_repstr = .03, lab = "*", roi = factor(1, levels = c("ahip","phip","mofc")))

ggplot(facat_sum, aes(x = factor(rep_type, levels = c("itemrep2","catrep")), 
                      y = m_repstr, 
                      color = factor(rep_type, levels = c("itemrep2","catrep")))) +
  geom_point(stat = "identity", size = 3) +
  geom_hline(yintercept = 0, color = "#343434", alpha = .5) +
  geom_errorbar(aes(ymin = m_repstr - se_repstr, ymax = m_repstr + se_repstr, color = rep_type),
                width = .2) +
  facet_wrap(~roi) +
  scale_color_paletteer_d(package = "rcartocolor", palette = "Vivid",
                         name = NULL, labels = c("Item", "Category")) +
  scale_y_continuous(limits = c(-.015, .035)) +
  geom_text(data = facat_annotate, label = "*") +
  theme_minimal(base_size = 16) +
  labs(y = "Cohen's d", x = " ",
       caption = "Figure 1. Strength of item and category representations") +
  theme(panel.grid.major.x = element_blank(),
        panel.grid.major.y = element_line(size = .3),
        panel.grid.minor.y = element_line(size = .3),
        legend.position = "none",
        plot.caption = element_text(hjust = -.35),
        axis.text.x = element_blank(),
        axis.text.y = element_text(color = "black"))

# So here is my recommendation for this plot. I think your plot may look good with points rather than boxes so I tried to change it to points. I think it looks good but it might just be personal preference here. I also tried to put the asterisk only for the Ant HIP roi plot but for some reason I could only get it to show up on an additional NA group. I have included a link below that shows various ways of people putting a single point. It appears that you have to create a data frame and include the coordinates for where you would want the point to show up. I got that far but I couldn't get it to show up on the right faceted plot.-JP

# https://stackoverflow.com/questions/11889625/annotating-text-on-individual-facet-in-ggplot2


```

### Figure 2. Behavioral Measures

+ Data: FACAT behavioral test measures
  ++ Measures for specific memory - recollection of full, first, and surname; recognition of old/new faces (corrected hit rate)
  ++ Measures for generalized memory - false alarms (recognizing new faces as old), categorization accuracy for new faces, increase in perceptual similarity of faces in the same category from pre to post training
+ Figure: Correlation matrix for the different task measures. 
+ Message: To show that test measures of specific and generalized/gist memory are not related. Also to show that categorization of new faces was more correlated with measures of memory specificity, suggesting participants might have used memory for the unique faces to categoryize new faces (exemplar strategy) rather than a generalized memory for the category (prototype strategy). 
+ Audience: NIH study section
+ Notes: 

```{r fig2_data}
behav <- facat %>%
  mutate(behav_cateffect = postpre_mSim_Relevant - postpre_mSim_Irrelevant) %>% 
  select(catACCnew, behav_cateffect, FA, 
         CHR, mFirstName, mSurname, mFullName)

zbehav <- scale(behav)

zbehav_corrmat_lwr <- cormat_lower(zbehav, diag = FALSE)
```

```{r fig2_v1}
# I like this because it's easier, but the flexibility in how the plot looks is not the same
# ggcorrplot(behav_cormat, method = "square", type = "lower",
#            hc.order = TRUE, lab = TRUE, lab_size = 3,
#            outline.color = "white", colors = c("#1A66E9","white","#E9481A"))

ggplot(zbehav_corrmat_lwr,
       aes(x = Var1, y = Var2, fill = value)) +
  geom_tile(color = "white", alpha = .9) +
  geom_text(aes(label = round(value,2)), size = 4) +
  scale_fill_gradientn(colors = c("#088200", "#00823d", "#00836B", "white", "#CC7400", "#cc9900", "#ccb400"), values = scales::rescale(c(-.5, -.3, -.1, 0, .1, .3, .5)), name = "Pearson \nCorrelation") +
  coord_fixed() +
  scale_y_discrete(labels = c("Categorization", "Perceptual Similarity", "False Alarms",
                              "Corrected Hit Rate", "First Name", "Surname")) +
  scale_x_discrete(labels = c("Perceptual Similarity", "False Alarms", "Corrected Hit Rate",
                              "First Name", "Surname", "Full Name")) +
  labs(caption = "Figure 2. Correlation matrix of task measures") +
  theme_minimal(base_size = 14) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        axis.title = element_blank(),
        legend.justification = "top", 
        plot.caption = element_text(hjust = 1.4, size = 14))
  
# I really enjoyed this plot for its use in examining correlations between variables. The only thing I changed and you are more than welcome to change it back is to have the colors represent effect sizes for correlations at .1, .3, and .5. Other than that, I liked this plot and think that it does an excellent job at showing how correlated some of your variables are. I was trying to get the scale to change on the side to where it would show exactly where the colors changed but I couldn't figure it out. Other than that, I like this plot. I did extend the colors from the original html color codes you had so I apologize for including uglier colors for the extreme values. -JP

```


### Figure 3. Representations and Behavior

+ Data: FACAT
+ Figure: scatter plot between strength of item/category representations in different brain regions and categorization performance
+ Features: color highlighting could be used to identify significant correlations
+ Message: To show that item representations in the VMPFC tracked categorization accuracy for new faces, again suggesting that participants were using an exemplar strategy  
+ Audience: NIH study section
+ Notes: 

#### Version 1: Scatterplot

```{r fig3_data}
facat_sub <- facat %>% 
  filter(subnum < 33) %>% 
  select(subnum, postpre_mSim_Relevant, postpre_mSim_Irrelevant,
         catACC, catACCnew, catACCold, 
         CHR, FA, mFirstName, mSurname, mFullName,
         contains("_train_")) %>% 
  mutate(behav_cateffect = postpre_mSim_Relevant - postpre_mSim_Irrelevant) %>% 
  gather(key = "tmp", value = "rep_str", 12:23) %>% 
  separate(col = "tmp", into = c("roi", NA, "rep_type")) %>%
  spread(key = "rep_type", value = "rep_str") %>%
  filter(roi != "hip")

```

```{r fig3_v1_stats, include=FALSE}
cat_corrs <- list()
item_corrs <- list()
for (i in 1:nrois) {
   tmp <- facat_sub %>% 
    filter(roi == rois[i]) 
   cat_corrs[[i]] <- cor.test(tmp$catACCnew, tmp$catrep)
   item_corrs[[i]] <- cor.test(tmp$catACCnew, tmp$itemrep2)
}

cat_regress <- list()
for (i in 1:nrois) {
  tmp <- facat_sub %>% 
    filter(roi == rois[i])
  cat_regress[[i]] <- lm(tmp$catACCnew ~ tmp$catrep + tmp$itemrep2)
}

map(cat_regress, summary)
```


```{r fig3_v1_data2}
facat_sub <- facat_sub %>% 
  select(-itemrep) %>% 
  gather(key = "rep_type", value = "rep_str", itemrep2:catrep)
```

```{r fig3_v1, fig.height=4.2, fig.width=10.5}
facat_sub$roi <- factor(facat_sub$roi, levels = c("ahip","phip","mofc"),
                        labels = c("Ant. HIP", "Post. HIP", "VMPFC"))

facat_sub %>% 
  ggplot(aes(x = rep_str, y = catACCnew, color = rep_type)) +
  geom_jitter(alpha = .8) +
  stat_smooth(geom = "line", se = FALSE, alpha = .3, method = "lm",
              size = 1) +
  geom_smooth(data = filter(facat_sub, roi == "VMPFC" & rep_type == "itemrep2"),
              se = FALSE, method = "lm", size = 1) +
  # stat_smooth(data = filter(facat_sub, roi == "VMPFC" & rep_type == "itemrep2"),
  # color = "gray70", method = "lm", se = FALSE,
  #             formula = y ~ x + I(x^2)) +
  facet_wrap(~roi) +
  theme_minimal(base_size = 16) +
  labs(x = "Strength of Representation \n(Cohen's D)",
       y = "Categorization Accuracy \n(new faces)",
       title = "Item representations in the VMPFC predict category generalization",
       caption = "Ant.HIP = Anterior Hippocampus \n Post.HIP = Posterior Hippocampus \n VMPFC = Ventromedial Prefrontal Cortex") +
  scale_color_paletteer_d(package = "rcartocolor", palette = "Vivid",
                         name = NULL, labels = c("Category Rep.", "Item Rep.")) +
  theme(legend.position = "bottom",
        plot.caption = element_text(hjust = -.3, size = 14),
        axis.text = element_text(color = "black"))

# I really like that using \n adds a space to the labs sections. I think this is great for when you want to include something in quotations like you have. In regard to the plots, I didn't think I could add any areas for improvement here. I did just change your points to jitter and I also added and commented out a way to look at quadratic associations. I know it doesn't help here much but I thought it might be helpful in the future. I do like how you used the alpha for stat_smooth lines so the audience's attention is focused on the last plot's line. I still do enjoy these colors. While I'm not sure if they are color blindness friendly, I think they would show up in presentations nicely or in a paper. -JP

```

#### Version 2: Betas

```{r fig3_v2_data, include=FALSE}
cat_regress <- lm(data = facat, 
                  catACCnew ~ ahip_train_itemrep2 + ahip_train_catrep +
                    phip_train_itemrep2 + phip_train_catrep +
                    mofc_train_itemrep2 + mofc_train_catrep)

summary(cat_regress)

mdl <- broom::tidy(cat_regress) %>% 
  mutate(ci_lower = estimate - (1.96)*std.error,
         ci_upper = estimate + (1.96)*std.error) %>% 
  filter(term != "(Intercept)")

mdl$term <- factor(mdl$term, levels = c("ahip_train_itemrep2", "ahip_train_catrep",
                                        "phip_train_itemrep2", "phip_train_catrep",
                                        "mofc_train_itemrep2", "mofc_train_catrep"))  
```

```{r fig3_v2}

term_color <- ifelse(mdl$p.value < .05, "#15C18C", "#6E6E6E") 

ggplot(mdl, aes(x = term, y = estimate)) +
  geom_hline(yintercept = 0, linetype = "dotdash", color = "gray", size = 1) +
  geom_errorbar(aes(ymin = ci_lower, ymax = ci_upper), width = .2,
                alpha = .9, color = "#6E6E6E") +
  geom_point(alpha = .8, color = "black") +
  geom_errorbar(data = filter(mdl, term == "mofc_train_itemrep2"),
                aes(ymin = ci_lower, ymax = ci_upper), 
                width = .2, color = "#15C18C") +
  geom_point(data = filter(mdl, term == "mofc_train_itemrep2"), 
             color = "#15C18C") +
  coord_flip() +
  theme_minimal(base_size = 14) +
  labs(x = NULL, y = "Beta Estimate",
       caption = "Figure 3. Item representations in the VMPFC predict category generalization") +
  scale_x_discrete(limits = rev(levels(mdl$term)),
                   labels = c("VMPFC Cat Rep.", "VMPFC Item Rep.",
                              "Post. HIP Cat Rep.", "Post. HIP Item Rep.",
                              "Ant. HIP Cat Rep.", "Ant. HIP Item Rep.")) +
  theme(axis.text.y = element_text(color = rev(term_color)),
        plot.caption = element_text(hjust = 1.5, size = 14))

# I enjoy the attention drawn to this graph through the use of color. An alternative could also be to use the gghighlight package that was taught in class but I assume it would get the same outcome. I do like the use of the horizontal line because I think it does an excellent job of drawing the audience's attention to the center of the plot. The only changes I made were from the dashed line to a dotdash line as I feel like it easier to see. I also changed the color of your points because I think it's a little easier for the audience to see the point for "Ant. HIP Item Rep." since it was close to the horizontal line and made it harder to distinguish them apart since all the colors are similar. Other than that, similar to all the other plots, I think this one does an excellent job at disseminating your findings. 
```


### Figure 4. Category Learning Model Fits

+ Data: DICAT 
+ Figure: Bar plot showing the average correlation between the model fits (exemplar v. prototype) and activity in each brain region. 
+ Features: scale color by each strategy (prototype v. exemplar), could highlight just prototype 
+ Message: Neural activity in certain brain regions track the type of strategy people use to make category judgements of new items.
+ Notes: I feel like this is incredibly similar to my figure 1. As I was creating the plots it kind of jumped out at me how I should present this (incidental v. feedback-based category learning). I'm thinking I might actually combine this figure with figure 1 to compare the results from two different tasks. 

```{r dicat}
dicat <- dicat %>% 
  select(subnum, contains("avg"))

dicat_tidy <- dicat %>% 
  gather(key = "tmp", value = "mbeta", -subnum) %>%
  separate(col = "tmp", into = c("mdl_fit","roi",NA))

dicat_sum <- groupsum(dicat_tidy, y = mbeta, g1 = mdl_fit, g2 = roi)
```

```{r dicat_stats, include=FALSE}
nvar <- length(dicat) - 1

test_betas <- list()
for (i in 1:nvar) {
  test_betas[[i]] <- t.test(dicat[i+1])
  test_betas[[i]]$data.name <- colnames(dicat)[i+1]
}
```

```{r fig4_v1}

ggplot(dicat_sum, aes(x = roi, y = mean, fill = mdl_fit)) +
  geom_hline(yintercept = 0, color = "#343434", alpha = .9) +
  geom_bar(stat = "identity", position = "dodge", color = "white") +
  geom_errorbar(aes(ymin = mean-se, ymax = mean+se),
                position = position_dodge(width = .9), 
                width = .2, color = "#343434") +
  scale_fill_OkabeIto(name = NULL,
                      labels = c("Exemplar Fits", "Prototype Fits")) +
  labs(y = "Parameter Estimates", x = NULL) +
  scale_x_discrete(labels = c("Ant. HIP", "Post. HIP", "VMPFC")) +
  theme_minimal(base_size = 16) +
  theme(legend.justification = "top")

# While similar I think this plot is just as interesting. I think it is stronger on its own but its up to you. The only change I made was putting the order of the regions of the brain in the same order as all your other plots. It might make your audience think the differences lie in the Post. HIP at first glance since that has been the order of all your other plots. Other than that, I like this one. -JP
  
```

```{r fig4_v2, include=FALSE}
ggplot(dicat_tidy, aes(x = roi, y = mbeta, color = mdl_fit, fill = mdl_fit)) +
  geom_bar(stat = "summary", fun.y = "mean", position = "dodge", alpha = .4) +
  geom_point(aes(x = roi, y = mbeta, color = mdl_fit),
    alpha = .5, position = position_dodge(width = .9)) +
  theme_minimal()

ggplot(dicat_tidy, aes(x = mdl_fit, y = mbeta, group = subnum, color = roi)) +
  geom_point() +
  geom_line(alpha = .5) +
  facet_wrap(~roi)
```


### Figure 5. Categorization Strategy

I have nothing interesting to show here yet, still playing around with the data and running analyses. 


```{r aepet}
aepet <- aepet

mdl_sum <- groupsum(aepet, y = PET_g, g1 = PET_cat_isprot)
mdl_sum$PET_cat_isprot <- factor(mdl_sum$PET_cat_isprot, 
                                 labels = c("exemplarist","prototypist"))

ggplot(mdl_sum, aes(x = PET_cat_isprot, y = mean, color = PET_cat_isprot)) +
  geom_point(stat = "identity") +
  geom_errorbar(aes(ymin = mean-se, ymax = mean+se), width = .2) +
  labs(x = "Dominant Categorization Strategy",
       y = "Mean Categorization Accuracy \n(new stimuli only)") +
  scale_fill_hc(guide = "none") +
  theme_minimal(base_size = 16) + 
  coord_flip()

# Similar to the other plot, I think it is possible to include this information as points rather than bar graphs. I think this is better at disseminating your findings as the bar plots were huge and taking up a large section of the plot without truly showing the differences between the two categories. I also just added a coord_flip() because I like seeing plots sideways. -JP
```

```{r, include=FALSE}
mdl <- lm(data = aepet, PET_g ~ PET_cat_exfit + PET_cat_protfit)
summary(mdl)

cor.test(aepet$PET_g, aepet$PET_cat_protadv)
cor.test(aepet$PET_g, aepet$PET_s)
cor.test(aepet$PET_g, aepet$MOT_s)
```

```{r}
aepet %>% 
  ggplot(aes(x = PET_cat_protadv, y = PET_g)) +
  geom_jitter(alpha = .3) +
  stat_smooth(color = "gray50", method = "lm", 
              se = FALSE, formula = y ~ x + I(x^2)) +
  geom_smooth(method = "lm", color = "#15C18C") +
  theme_minimal() +
  labs(x = "Prototype Advantage \n(i.e. prototype model fit better)",
       y = "Category Generalization Accuracy")

# Since this is in regard to model fits, I think it might be helpful to include curvilinear fits. I'm not sure about neuro-related research and looking at curvilinear associations but I think it is interesting if there is a theroetical basis for investigating if a model is curvilinear. Other than that, I think this is a great plot and a good addition to everything else that you included in this file. Great work and look forward to your blog/website for this project. -JP
```


